% Copyright 2011, Piotr Jakubowski

\chapter{Research}
  The aim of this chapter is to present the environment in which the plugin will be developed. I will present, firstly, the features of Ruby language and then what is Ruby on Rails, how it is constructed and how we can enrich it with plugins. In the end I will take a look into possibilities of metaprogramming in Ruby and Rails.
  
  \section{Ruby}
  Here you will find information abut Ruby language and especially those features that create the power of Ruby and distinguish it from other programming languages.
    \subsection{Basic information}
      \subsubsection{History}
      Ruby language has been created by Yukihiro Matsumoto also known as "Matz" for english speaking programmers. It firstly appeared publicly in 1995 with version 0.95 in order to reach version 1.0 a year later. As of the time of writing this document, the latest release of Ruby is of branch 1.9 (specifically 1.9.2), but the branch 2.0 with brand new exciting features is emerging on the horizon.
    
      While creating Ruby Matz stated that he focused on developing a language that would be programmer-friendly and that would make the barrier between idea and putting this idea in code as low as possible. The most famous quote of Matz fully describes the philosophy of the language:
    
      \begin{quote}
        Ruby is designed to make programmers happy
      \end{quote}
        
      \subsubsection{Implementation}
      The official implementation of Ruby has been written in C language. As no particular standard has been developed for the language, the official implementation is a reference for all other vendors. Nonetheless, there is plenty of other implementations including one operating in JVM(JRuby\footnote{http://www.jruby.org/}), .NET framework(IronRuby\footnote{http://ironruby.codeplex.com/}) and Objective-C runtime(MacRuby\footnote{http://www.macruby.org/}).
    
    \subsection{Features}
      \subsubsection{Basic features}
      It is worth mentioning few basic features of Ruby language that would make the further part of this chapter clearer.
      \begin{itemize}
        \item Ruby is scripting language - statements are executed as provided and there is no special \texttt{main} function
    	  \item Ruby is dynamically typed
    	  \item Method invocations do not need to include parentheses
    	\end{itemize}
    	
      \subsubsection{Everything is an object}
      Ruby is an object oriented language. In fact, every value in Ruby is a object. Even such "primitive" values as integers, true, false or nil (which is Ruby's NULL). Thanks to that, Ruby promotes and enforce object oriented programming. 
      
      Moreover, you can have methods on integers or other primitive values which makes the code much more readable, shorter and enjoyable. For example instead of having Java-like:
      
      \lstinputlisting[language=java]{code/chapter03/objects.java}
      
      You can have much more readable:
      
      \lstinputlisting[language=ruby]{code/chapter03/objects.rb}
      
      Ruby standard library does not include \texttt{to\_roman} function for integers, but we could easily add one by using next described feature.
      
      \subsubsection{All classes are open}
      Ruby gives us freedom to change already declared classes. Moreover, it does not mean that we can change only the classes that we defined. We can change even classes defined by Ruby standard library. Therefore we can open for example class Integer that represents all integer values in Ruby code and add method \texttt{to\_roman}:
      
      \lstinputlisting[language=ruby]{code/chapter03/openclasses.rb}
      
      Now, we would be able to have following line in our code:
      
      \lstinputlisting[language=ruby]{code/chapter03/openclasses2.rb}
      
      \subsubsection{Power of blocks}
      This is one of the most exciting features of Ruby language. Block is a fragment of code that can be passed to a function and the function may call this code anywhere inside its body. 
      
      Block in ruby can be denoted in two ways:
      
      \lstinputlisting[language=ruby]{code/chapter03/blocks01.rb}
      
      Usually the first way is used for blocks that span throughout multiple lines, while the second way is used for single line blocks. 
      
      Blocks can be used in number of ways. The most popular are iterators:
      
      \lstinputlisting[language=ruby]{code/chapter03/blocks02.rb}
      
      Above code would result in printing 1, 2, 3 to the output. Statement \texttt{do |i|} means that this block expects one argument called \texttt{i} just like regular functions do.
      
      In order to understand how blocks operate, let me present simple example:
      
      \lstinputlisting[language=ruby]{code/chapter03/blocks03.rb}
      
      The above example adds the \texttt{each\_nested} method to an array that allows us to perform operations on every element of multidimensional matrix. It iterates over every element of an array. If given element is again an array then it recursively calls \texttt{each\_nested} on it and if it is other element then it passes the control to a block along with the element as a parameter.
      
      As we can see appropriate use of blocks may be very useful and can make code much shorter and more readable
      
    \subsection{Summary}
    This chapter demonstrated basic features of Ruby language. In order to get more information on this language I forward the reader to the official website\footnote{http://ruby-lang.org/} or to some books from bibliography\cite{ruby01}.
    
  \section{RubyGems}
    \subsection{Basic Information}
  RubyGems\footnote{http://rubygems.org/} is the packaging system to distribute libraries for Ruby language. The packages and libraries are called gems and console front-end script is called (not surprisingly) \texttt{gem} (distributed along with ruby 1.9, for ruby 1.8 needs to be installed separately). With the help of RubyGems Ruby program can be easily enhanced our with variety of functionalities (for instance serializing/deserializing JSON, handling weekends and holidays in Dates, authentication system etc). 
  
  In order to use particular gem first it has to be installed with \texttt{gem} command 
  \begin{lstlisting}
    gem install some_gem
  \end{lstlisting}
  
  And then it has to be added to application by first requiring rubygems itself and then requiring the gem.
  
  \begin{lstlisting}
    require 'rubygems'
    require 'some_gem'
  \end{lstlisting}
  
  With those few lines of code it is possible to add really extensive features to Ruby programs, as the community gets bigger and bigger and, moreover, its members are willing to share the code.
  
    \subsection{Dependency management}
  Of course, every gem can use other gems in order to build up on their functionalities. However, in order for those gem to function properly all their dependencies. Moreover, different version of the same gem can differ not only in the implementation, but also in their API. Therefore gems need to keep track of their versions. Surely, it would be totally inefficient if users would have to take care of this on their own. This is why RubyGems take care of it for us.
  
  Every gem is described by its gemspec (gem specification) file. Gemspec is simply Ruby code that defines particular gem. Below is presented gemspec for Ruby on Rails:
  
  \lstinputlisting[language=ruby]{code/chapter03/rails.gemspec}
  
  As seen above, gemspec is pretty straightforward. Along with other attributes, developer can specify dependencies of his gem. Then, when user installs it RubyGems traverses the list of dependencies and installs all of them (of course in the meantime traversing their gemspecs and installing theirs dependencies, which results in the entire tree of dependency). That way, RubyGems becomes really efficient tool to handle libraries. 
  
  \section{Ruby on Rails}
    \subsection{Introduction}
  Ruby on Rails is a comprehensive framework for web applications development. It has been created by David Heinemeier Hansson in 2004 and a year later he earned Google and O'Reilly's Hacker of the Year award. Its main characteristic is promoting convention over configuration, what enables developers to quickly build even complex web apps without the need to write lengthy XML configuration files. It utilizes Model-View-Controller pattern, which happens to be very good choice for web applications and helps produce clean and maintainable code.
  
  Ruby on Rails is distributed as a gem for Ruby language. After just one command (provided that Ruby and RubyGems are installed on the system):
  
  \begin{lstlisting}
    gem install rails
  \end{lstlisting}
  
  the user has entire environment for creating web applications (along with development server).
  
  Among others, Ruby on Rails has two main parts:
  \begin{description}
    \item[ActiveRecord] Object-relational mapper
    \item[ActionPack] Framework for handling requests and rendering responses
  \end{description}
  
  This two parts mainly create the power of the framework and will be described in more details later.
  
    \subsubsection{Structure of the application}
    As mentioned before, Ruby on Rails takes its power and agility from "convention over configuration" rule. Therefore, Ruby on Rails applications have very well defined directory structure. The user can easily generate one for his application by using following command:
    \begin{lstlisting}
      rails new name_of_application
    \end{lstlisting}
    
    Or in case the user is using older (below 3.0) version of Rails:
    
    \begin{lstlisting}
      rails name_of_application
    \end{lstlisting}
    
    This command generates the entire directory tree that is used in Rails application. Most important directories are described below.

    \begin{description}
      \item[app] Contains main code of applications. It is divided into models, controllers and views subfolders
      \item[config] Holds configuration files
      \item[db] Holds files that regard database
      \item[lib] Contains files that create environment for the application but are considered as strictly connected to business logic of application
      \item[public] Place for assets of the application: static pages, stylesheets, javascript files - anything that can be statically served.
    \end{description}

    Thanks to such well-defined structure, Ruby on Rails knows exactly where to look for particular parts of the system. Moreover, it does not need to go through all the files, but rather load them on demand.

    \subsection{ActiveRecord}
      \subsubsection{Introduction}
      ActiveRecord is one of Ruby's Relational Object Mappers, but has been developed as
      part of Ruby on Rails framework. As whole framework, ActiveRecord philosophy is convention
      over configuration.

      Usually, models in Rails applications are stored in app/models directory. They are classes that
      derive from ActiveRecord::Base class. And, in fact, this is enough to handle simple models.
      It is possible to create persistent business models with great ease.

      \lstinputlisting[language=ruby]{code/chapter03/activerecord01.rb}

      Above example depicts the most simple model. Of course, the user needs to set up configuration
      for database connection. Usually it is done in config/database.yml file while in Ruby on Rails project
      or provide paramaters directly to ActiveRecord (if not used in Rails project).

      \lstinputlisting[language=ruby]{code/chapter03/activerecord02.rb}

      \subsubsection{Basic Model}
      Coming back to the code example of model:

      \lstinputlisting[language=ruby]{code/chapter03/activerecord01.rb}

      This will map automatically to items table in the database. It is not necessary to specify
      what kind of columns, the table has. All attributes of the model are automatically
      drawn from the table and (thanks to dynamic nature of ruby) translated into appropriate
      accessor methods for Ruby class.

      So assuming the table has been created with following sql statement

       \lstinputlisting[language=ruby]{code/chapter03/activerecord01.rb}

      Then with the ruby code presented above we get following functionallity:

       \lstinputlisting[language=ruby]{code/chapter03/activerecord04.rb}

      Moreover, it is possible to search for records by forming (even very complex) queries:

      \lstinputlisting[language=ruby]{code/chapter03/activerecord05.rb}

      \subsubsection{Associations}
      As every Object Relational Mapper, ActiveRecord enables developers to specify relations
      between models. The only thing that needs to be done in order to handle associations well
      is add the foreign key column to the database and describe the relation in the model by
      using one of three class methods:
        \begin{itemize}
          \item{has\_many}
          \item{has\_one}
          \item{belongs\_to}
        \end{itemize}
      This is shown in the example:

      \lstinputlisting[language=ruby]{code/chapter03/activerecord06.rb}

      \subsubsection{Validations}
      As models usually keep great deal of application's business logic there is another feature
      that is very useful - validations. It is possible to define rules against which the
      model will be tested. If some of those conditions would not be fulfilled, the record would
      not be saved and ActiveRecord would store Errors object that gathers information on
      unsatisfied validations.

      \lstinputlisting[language=ruby]{code/chapter03/activerecord07.rb}

      Functionality presented above is just a small subset of possibilities. It is possible
      to specify when given validations should be executed (during create or update of the record)
      or specify conditions that turn off validation (eg. do not check presence of one attribute
      if the second one is true). Moreover, except for standard \texttt{validates\_*} functions,
      the user can specify his own validations by passing the names of methods that would be
      executed during validations to \texttt{validate} class method.

      \subsubsection{Migrations}
      Among others, ActiveRecord provides another feature that makes the life of developers
      just a little less painful - Migrations. They are used to handle changes to the database,
      and as an extension - to structure of Models. 

      So, instead of creating and altering tables from SQL we could use a migration that looks
      like this:

      \lstinputlisting[language=ruby]{code/chapter03/activerecord08.rb}

      Rails provide tools that make work with migrations like generators that help create migration
      files or tools for running migrations (\texttt{rake db:migrate}). Moreover, in the applicaiton
      database Rails would create a table called \texttt{schema\_migrations} that keeps track of 
      which migrations have been applied to given database. This is useful, when working on some
      project with fellow developers - if one of them adds a migration and then the other developer
      pulls it from the version control system, he can run \texttt{rake db:migrate} and Rails
      would run only new migrations.

      \subsubsection{Summary}
      As stated in the introduction, it is clear to observe, that ActiveRecord provides great deal
      of features that surely makes the rapid development of applications a breeze. Moreover, thanks
      to Ruby's nature it all is packed in very easy and accessible API.

      The intent of this chapter was to make a short introduction to ActiveRecord and it did not cover
      all of interesting and useful features of ActiveRecord. There is plenty of other things such as
      callbacks, observers, named\_scopes that are worth noting, but in order to keep
      this document as focused on the topic as possible they are omitted here. In order to get more
      information, the reader can refer to ActiveRecord documentation.
      
    \subsection{ActionPack}
      \subsubsection{Introduction}
      As said before, Ruby on Rails is build on the Model-View-Controller pattern. Model part is
      handled by ActiveRecord described in previous chapter. ActionPack takes care of two remaining
      parts - View and Controller. So, in general ActionPack is divided into two parts:
      \begin{itemize}
        \item{ActionController}
        \item{ActionView}
      \end{itemize}
      It is easy to guess which part is responsible for which part of MVC pattern.
      
      \subsubsection{From request to response}
      In very short, when Ruby on Rails application gets a request, first, it parses it in order 
      to translate the request into form that is better to use programatically. Then, based on the
      routing rules set in your application Rails determines which action from which controller
      to call. Then it creates an instance of particular controller and calls method for 
      target action and passes the parameters that came along with the request.
      
      Then the particular controller takes over. Its responsibility is to understand the data
      provided with request, conduct some actions based on them and render appropriate 
      response. The response is usually rendered by using some template that is filled with data
      provided by the controller.
      
      \subsubsection{ActionController}
      Controllers in Ruby on Rails are just subclasses of ApplicationController, which inherits
      from ActionController::Base. Actions are defined by just simply adding instance methods
      to controller class.
      
      The most basic controller for blog posts could look like this:

      \lstinputlisting[language=ruby]{code/chapter03/actioncontroller01.rb}
      
      The controller presented above has 5 basic CRUD actions. It can be observed, that controller
      assigns instance variables - this is in order to be able to read those variable in the view.
      Moreover, some of the actions of this controller do not have \texttt{render} command which would 
      cause the chosen template to be processed. This is correct, because Rails by default 
      (if no \texttt{render} has been specified) looks for template in default path - in this
      case it would be \texttt{app/views/posts/action\_name}, so for example for \texttt{index}
      action it would look for \texttt{app/views/posts/index.html.erb}
      
      \subsubsection{ActionView}
      ActionView is Ruby on Rails' way to generate html (and not only - xml and others can also be
      produced) views. ActionView by default provides three ways to generate views:
      
      \begin{description}
        \item[erb] ERB (Embedded Ruby) templates -  mix of html and embedded ruby code
        \item[builder] builder for XML documents
        \item[rjs] generator of javascript actions
      \end{description}
      
      The most common are ERB templates. They enable developers to easily define
      HTML views. The example of ERB template could look like this:
      
      \lstinputlisting[language=ruby]{code/chapter03/actionview01.erb}
      
      As said in previous chapter, the instance variables of controller are available in the view.
      That is why it is possible to use \texttt{@post} variable and thanks to this it is possible to 
      extract all the logic into controller and have clean views.
      
      \subsubsection{Summary}
      This chapter was just the introduction into ActionPack in order to give an overview of how 
      it works. It does not cover more advanced features as filters in controllers or layouts
      and partials in views.
      
  \section{Metaprogramming}
  Metaprogramming is programming of code that its input and output is other code. It is used to 
  manipulate the behavior of the program during runtime depending on other parts of the application.

  Thanks to dynamic nature of Ruby it provides comprehensive reflection API. It allows developer
  to modify classes, list or add methods to classes or even particular objects.
    
    \subsection{Metaprogramming in Ruby}
      The way Ruby language is constructed makes it really easy to apply metaprogramming to. First of   
      all, it is a dynamic language which means that things like adding new code or extending objects 
      can be done during run time. It is kind of a direct result of the fact that
      Ruby is an interpreted language. Anyway, this single characteristic makes it very useful in
      terms of metaprogramming. What is more, as everything in Ruby is an object so are classes 
      (they are instances of class Class). Therefore, we can act on them as we would do on objects, 
      what sounds like even bigger improvement on metaprogramming.
    
      As said before, developer can define methods during runtime. Moreover, it is possible to reopen 
      class definitions multiple times and add methods. The following example depicts this
    
      \lstinputlisting[language=ruby]{code/chapter03/meta01.rb}
    
      The class \texttt{MyClass} is first created, but without any methods. Then the object of this
      class is created. So far it is pretty standard, but next the user is asked for an input. Then
      depending on this input the class \texttt{MyClass} is reopened and method \texttt{what?} is
      defined with implementation depending on user input. Then the method is called on previously
      created object. After running this program the user would be able to observe that depending
      on what is his input the appropriate method definition is applied.
    
      The example above shows the dynamic nature of Ruby. But this is only the beginning. Thanks
      to so-called Singleton Classes in Ruby it is possible to define methods on particular objects.
      So it is possible to rewrite the example in such way:
    
      \lstinputlisting[language=ruby]{code/chapter03/meta02.rb}
    
      This time \texttt{what?} method has been defined for \texttt{object}. If developer would create
      another instance of class \texttt{MyClass} it would not have \texttt{what?} method. As soon as
      method has been defined for particular object the class of this object has been changed to 
      singleton class - a subclass of \texttt{MyClass} - and it keeps all the methods defined for the 
      \texttt{object}. If digged deeper into Ruby internals it would be 
      possible to see that such class has "singleton" flag set to true. This is an indication for Ruby
      that this class is only a helper class and would not be seen in example in \texttt{object.class}
      call (this method would still yield \texttt{MyClass}).
      
      Those examples show the dynamic nature of Ruby language. In addition to features presented above, 
      Ruby provides a lot more. For example, it gives the ability of evaluating a string as Ruby code.
      This functionality is provided by following methods:
        
        \begin{itemize}
          \item{eval}
          \item{instance\_eval}
          \item{class\_eval}
        \end{itemize}

      The difference between those methods is the context the string is going to be evaluated in. 
      \texttt{Eval} evaluates code in current context (or in the context passed to eval by providing
      Binding object as second argument), \texttt{instance\_eval} is called on some object and evaluates
      the code in context of the receiver of the method while \texttt{class\_eval} can be called on a 
      Class or Module and evaluates the code in context of given Class/Module.
      
      Moreover, the developer can find more methods that make the reflection API complete:
      
        \begin{itemize}
          \item{methods}
          \item{protected\_methods}
          \item{private\_methods}
          \item{define\_method}
          \item{remove\_method}
          \item{undef\_method}
        \end{itemize}
      
      Names of the methods are self explanatory. The difference between \texttt{remove\_method}  and
      \texttt{undef\_method} is that \texttt{remove\_method} removes implementation from particular module 
      but Ruby will still traverse the inheritance tree to find implementations of given method in parent
      classes. \texttt{Undef\_method},on the other hand, would prevent the module from responding to given
      method at all.
      
    \subsection{Metaprogramming in Rails}
      As Rails is a framework for Ruby it is sane to think that it would provide the same level of
      dynamism and same level of metaprogramming capabilities. And this assumption turns out to be 
      valid. First of all, developers can make use of all Ruby metaprogramming capabilities. But Ruby on 
      Rails introduces some new capabilities on top of Ruby plus adds functionality that is specific for
      web apps, object relational mapping etc.
      
      One of most useful methods added in Rails is constantize methods. It takes a string as an argument and
      returns a constant that is found in the environment under this name. Therefore, as Ruby classes
      are just constants that point to particular objects of Class we may get the classes by their names
      in string as presented in the example:
      
      \lstinputlisting[language=ruby]{code/chapter03/meta03.rb}
      
      Such usage of constantize may be very useful, especially in kind of application this thesis
      presents, as it would allow to pass the names of entities around as strings and dynamically retrieve
      classes in the code.
      
      In addition, ActiveRecord provides comprehensive reflection API. Developer can get the list of all
      columns the entity has in database by calling \texttt{columns} method on the entity class. That method 
      returns list of subclasses of \texttt{ActiveRecord::ConnectionAdapters::Column} which provides
      information on name, type and other metadata of the column.
      
      In order to fully make use of this API and effectively manage entities there is another method 
      that may be helpful. Developer can call \texttt{reflect\_on\_all\_associations} which would 
      return the array of objects representing all associations (\texttt{has\_many}, 
      \texttt{belongs\_to}, \texttt{has\_one}) the 
      particular entity has. Moreover, if you pass association type as an argument it would return only
      associations of this type.
      
      Features shown above would definitely make writing the targeted plugin much easier and enjoyable.
      
    
      
      
      
      
        
      
      
      
      